---
title: "Mc"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  message = FALSE, 
  warning = FALSE,
  collapse = TRUE,
  error = TRUE,
  comment = "#>"
)
```

```{r}
library(FLCore)
library(FLXSA)
library(FLasher)
library(ggplotFL)
library(patchwork)
library(tidyverse)
# remotes::install_github("einarhjorleifsson/fishvice", dependencies = FALSE)
library(fishvice)
```

#### little helper functions

```{r}
# a retro functions
lh_retro <- function(end.year, fls, ind, cnt) {
  window(fls, end = end.year) + 
    FLXSA(window(fls, end = end.year), 
          window(ind, end = end.year), 
          cnt)
}
# get the reference biomass next year
lh_1year_projection <- function(fls) {
  # the mean recruitment model
  sr <- as.FLSR(fls, model="geomean")
  params(sr)['a',] <- exp(mean(log(rec(fls))))
  # use same weights and Fay (gets muliplied by Flevel) as in the terminal year
  proj <- stf(fls, nyears = 1, wts.nyears = 1)
  last.year <- range(fls)["maxyear"][[1]]
  fsq <- mean(fbar(fls)[,as.character(last.year)])
  ctrl <- 
    fwdControl(data.frame(year = last.year + 1, 
                          quant = "f", 
                          value = fsq))
  fwd(proj, control = ctrl, sr = sr)
}
# a compilation function (here only biomass, B4+)
lh_retro_B4p <- function(ryears, fls, ind, cnt) {
  retro <- map(ryears, lh_retro, fls, ind, cnt)
  # add one year projection to get the biomass in the ass year
  #  note the smb index in the ass year is NOT used in the fitting, need to check
  if(projections) retro <- map(retro, lh_1year_projection)
  names(retro) <- ryears
  bio <- 
    map(retro, computeStock) %>% 
    map(as.data.frame) %>% 
    map(as_tibble) %>% 
    bind_rows(.id = "assyear") %>% 
    mutate(assyear = as.integer(assyear)) %>% 
    ungroup()
  bio.last <- 
    bio %>% 
    filter(assyear == max(assyear)) %>% 
    select(year, bio = data)
  bio <- 
    bio %>% 
    left_join(bio.last) %>% 
    mutate(r = log(data / bio))
  # just so no need to alter code below
  return(list(bio = bio))
}
# the plot
lh_ggplot <- function(res, last.point, bio.spaly) {
  p1 <- 
    res %>% 
    ggplot() +
    geom_line(data = bio.spaly,
              aes(year, bio),
              colour = "yellow",
              lwd = 1) +
    geom_line(aes(year, data, group = assyear)) +
    geom_point(data = last.point, colour = "red",
               aes(year, data)) +
    facet_wrap(~ run) +
    labs(x = NULL, y = "B4+")
  p2 <- 
    res %>% 
    ggplot(aes(year, r, group = assyear)) +
    geom_line() +
    geom_point(data = last.point,
               colour = "red") +
    facet_wrap(~ run) +
    labs(x = NULL, y = "contemporaneous\n vs current")
  return(p1 + p2 + plot_layout(ncol = 1))
}
# doing retro statistic using the last ~5 years of the assessment 
#  is plain stupid, even though ices does so
lh_retro_performance <- function(d) {
  left_join(d %>% 
              group_by(run) %>%  summarise(bias98.21 = mean(r), cv98.21 = sd(r)),
            d %>% filter(year %in% 1998:2017) %>% 
              group_by(run) %>%  summarise(bias98.17 = mean(r), cv98.17 = sd(r)))
}
```


### get data

```{r inputs}
YRS <- 1985:2021   # it is only fair to use 1994 onwards, the time the
                   #  estimates start, but here just trying to repeat the paper
fil <- "/u2/reikn/Tac/2022/01-cod/data-raw/Supplementary S3.xlsx"
shits <- readxl::excel_sheets(fil)
mc <- 
  expand_grid(year = YRS,
              age = 1:14) %>% 
  left_join(bind_rows(readxl::read_excel(fil, sheet = shits[1]) %>% mutate(sur = "smb"),
                      readxl::read_excel(fil, sheet = shits[2]) %>% mutate(sur = "smh"))) %>% 
  spread(sur, Mc) %>% 
  mutate(mc = case_when(year < 1994 ~ 0.2,
                        is.na(smh) ~ smb,
                        TRUE ~ (smb + smh) / 2)) %>% 
  select(year, age, mc) %>% 
  arrange(year, age) %>%
  group_by(year) %>% 
  # assume same mc in all year after last datayear
  fill(mc, .direction = "downup") %>% 
  # 10+ same as 9
  group_by(age) %>% 
  fill(mc) %>% 
  ungroup() %>% 
  filter(year %in% YRS)
rbx <- fishvice::mup_rbx("/u2/reikn/Tac/2022/01-cod/ass/mup/smx", scale = 1000)
bio.spaly <- rbx$rby %>% filter(year %in% min(YRS):2022) %>% select(year, bio)
# have not figured out how to set xsa up with survey data in year with no catch data
#  besides of course the old-way, but can not be bothered with that now
rbx$rby <-  rbx$rby %>%  filter(year %in% YRS)
rbx$rbya <- rbx$rbya %>% filter(year %in% YRS)
# set maturity and stock weights such that SSB is actually B4+
rbx$rbya <-
  rbx$rbya %>% 
  mutate(mat = replace_na(mat, 0),
         mat = ifelse(age >= 4, 1, 0),
         ssbW = sW,                    # 2022-05-28 from cW to sW as in paper
         ssbW = replace_na(ssbW, 0))
rbx.mc <- rbx
# replace 0.2 with mc
rbx.mc$rbya <-
  rbx.mc$rbya %>% 
  select(-m) %>% 
  left_join(mc %>% rename(m = mc))
```

### flr setup

```{r xsasetup}
# pf and pm set to 0, so "SSB" = B4+ is based on stock in numbers in start of
#  year
# issue: age 14 should not be a plus group, for relative comparisons
#  should though be ok
fls <-    rbx %>%    rbx_to_flstock(pf = 0, pm = 0, fage = c(5, 10))
# range(fls)["plusgroup"] <- NA
fls.mc <- rbx.mc %>% rbx_to_flstock(pf = 0, pm = 0, fage = c(5, 10))
# range(fls.mc)["plusgroup"] <- NA
ind.smb <- 
  rbx_to_flindex(rbx, "U1", time = c(3/12, 3.5/12)) %>% trim(year = YRS)
# range(ind.smb)["plusgroup"] <- NA
# ind.smh <- 
#  rbx_to_flindex(rbx, "U2", time = c(10/12, 10.5/12)) %>% 
#  trim(year = 1996:2021)
# only spring survey used in paper
ind <- FLIndices(ind.smb)
```

### xsa retrospectives

#### controls

```{r xsaruns}
# power or no power, that is the question
cntrp0 <- FLXSA.control(maxit = 70, rage = 0, qage = 14)  # i would set qage to
cntrp5 <- FLXSA.control(maxit = 70, rage = 5, qage = 14)  #   age 9 or 10
```

#### run retro

```{r}
projections <- FALSE
ryears <- 1998:max(YRS)
retp0 <-    lh_retro_B4p(ryears, fls,    ind, cntrp0)
retp0.mc <- lh_retro_B4p(ryears, fls.mc, ind, cntrp0)
retp5 <-    lh_retro_B4p(ryears, fls,    ind, cntrp5)
retp5.mc <- lh_retro_B4p(ryears, fls.mc, ind, cntrp5)
```

### results

clarification on runs:
```
~run,    ~comment,
p002     No power,       M = 0.20
p0mc     No power,       M = mc
p502     Power ages 1:5, M = 0.20
p5mc     Power ages 1:5, M = mc
```

```{r}
res <- bind_rows(retp0$bio    %>% mutate(run = "p002"),
                 retp0.mc$bio %>% mutate(run = "p0mc"),
                 retp5$bio    %>% mutate(run = "p502"),
                 retp5.mc$bio %>% mutate(run = "p5mc"))
last.point <- 
  res %>% 
  filter(year == assyear + projections)
```

```{r, fig.height = 8, fig.cap = "Retrospective performance of 4 different model setups. See table above for further details on settings."}
lh_ggplot(res, last.point, bio.spaly)
```

#### perfomance measure

```{r}
last.point %>% 
  lh_retro_performance() %>% 
  knitr::kable(caption = "Retrospective performance measure of 4 different model setups. See table above for further details on settings.",
               digits = 3)
```

### conclusions on the retrospectives

* The variable M (p0mc) show better retrospective performance than constant M (p002) if no power is assumed.
* Both variable and constant M show improved retrospectives if a power up to and including age 5 is assumed.
* When power is assumed, the retrospective performance of constant M (p502) outperformes the variable M (p5mc).

There may though be more under the hood in the variable M diagnostics that may warrant further investigation. Something beyond that done in the Mc article.

### sanity check - the mc input

```{r sanitycheck}
m(fls.mc) %>% trim(year = 1993:2021, age = 1:10) %>% round(2)
```

### results of retro xsa with 1 year projections

... still not using the survey in the assessment year

```{r, echo = FALSE, fig.height = 8, fig.cap = "Retrospective performance of 4 different model setups. See table above for further details on settings."}

projections <- TRUE
ryears <- 1998:max(YRS)
retp0 <-    lh_retro_B4p(ryears, fls,    ind, cntrp0)
retp0.mc <- lh_retro_B4p(ryears, fls.mc, ind, cntrp0)
retp5 <-    lh_retro_B4p(ryears, fls,    ind, cntrp5)
retp5.mc <- lh_retro_B4p(ryears, fls.mc, ind, cntrp5)
res <- bind_rows(retp0$bio    %>% mutate(run = "p002"),
                 retp0.mc$bio %>% mutate(run = "p0mc"),
                 retp5$bio    %>% mutate(run = "p502"),
                 retp5.mc$bio %>% mutate(run = "p5mc"))
last.point <- 
  res %>% 
  filter(year == assyear + projections)
lh_ggplot(res, last.point, bio.spaly)
last.point %>% 
  lh_retro_performance() %>% 
  knitr::kable(caption = "Retrospective performance measure of 4 different model setups. See table above for further details on settings.",
               digits = 3)
```


### change log

```
2022-05-28:
* changed basis of reference biomass from cW to sW, reflecting what i think was 
   done in the paper. B4+ is actually based on catch weights. That basis
   is of course wrong, but it is linked to the 0.20 multiplier in the HCR that
   also includes a stabilizer.
   Whatever the case, for some reasons XSA gives a much higher biomass when
   using catch weights than muppet, or for that matter sam. Most likely related
   to mishandling of the plusgroup. But these details are not an issue in this 
   doodle.
* added one year prediction, i.e. into the advise year, to reflect what was
   actually done in the paper. note though that not using the survey index in 
   the advice year is still an issue (can not be done in XSA, i think).
* also simplified code, but that alone does not affecting previous versions 
   of this document
* added different periods for estimating retrospective performance. not yet 
   though repeating the 2 five year peels in the paper because that is not 
   kosher.
2022-05-27
  * converted lh_function, now used purrr::map rather than lapply
  * added window to the index as well, just for clarity
  * neither affects the results nor the conclusion
  * added conclusions and some minor additional text edits
2022-05-26 
  * first version - bb & co notified via email
```
